### 2020-09-19 response.tex

\revtext{I have one main concern, and this concerns the model. First a very general model is discussed, then a very special version is mentioned (mid p11) in which the latent period and incubation periods are always identical. For this model the General interval and Serial interval have exactly the same distribution. My concern is whether the stated results are valid only for this special case or if they apply also in general. I doubt the results hold in general, but at least think arguments for it are missing.}

Thank you for pointing this out. The special version is only presented in the supplementary material. We do not assume this in the main text. We have made this distinction clear:

``For example, the Susceptible-Exposed-Infected-Recovered model, under the additional assumption that the incubation and exposed periods are equivalent (i.e. that onset of symptoms and infectiousness occur simultaneously), provides a special case where the forward serial- and generation-intervals follow the same distributions during the exponential growth phase because (i) infected individuals can only transmit after symptom onset and (ii) the time between symptom onset and infection is independent of the incubation period of an infector (see Supplementary Materials).
Throughout the paper, we allow for presymptomatic transmission to match the transmission dynamics of COVID-19.''

Figure 2 further shows that the initial forward serial-interval distribution is different from the intrinsic generation-interval distribution but gives identical \Ro estimates given $r$ using parameter estimates for COVID-19.

@@

\revtext{The methods and results sections do excellent jobs explaining the source of confusion and biases, and how to overcome them. The discussion section synthesizes the insights into suggestions and good practices. However, I miss examples of how far off - if off at all - criticized application papers are when estimating the reproduction number.}

\revtext{The authors can improve their manuscript by responding to the following issues:}

\revtext{What is the manuscript about - generation and serial intervals or the reproduction number R? The title and abstract suggest the intervals and the introduction the reproduction number. I think opening with the well-known reproduction number will give the most transparent and coherent story.}

It is about both. Generation and serial intervals are key quantities in describing epidemic dynamics; their differences have long been misunderstood and our paper provides a theoretical basis for linking them properly. These intervals also play critical role in estimating the reproduction number. We have made major revisions to the writing of the manuscript based on the suggestions to provide a transparent and coherent story.

@@

\revtext{The reproduction number is not put in a nuanced context. The presentation of the reproduction number as "one of the most important characteristics of an emerging epidemic ... [definition] ... [that] allows us to predict the extent to which an infection will spread in the population, and the amount of intervention necessary to eliminate it" that then leads directly into how it is estimated leaves no room for its shortcomings. A more nuanced picture includes its inability to, for example, account for individual variation, including superspreaders and spatial variations in regions.}

We have added the following limiations of the reproduction number:

``Since the reproduction number represents an average \citep{diekmann1990definition, anderson1991infectious}, it does not capture heterogeneity among individuals or across space.
The reproduction number also does not provide any information about the time scale of disease transmission.''

@@

\revtext{The introduction presents previous methodological work on the R-r relationship based on generation and serial intervals and recent applications that confuse the two intervals. The presentation goes back and forth between the two and does not clarify the shortcomings of each work or application. My suggestion for a stronger introduction: A funnel focusing on the methods that ends with the knowledge gap "In contexts where the distributions are expected to be different, current theory has no explanation for how these differing distributions could provide identical estimates of R from r" (and further specified by the gaps indicated at the end of section 2.1 and beginning of section 2.2) followed by concrete examples of how off R estimates are in applications to substantiate the problem.}

We have rewritten a large proportion of the introduction section based on these suggestions.

### 2020-09-19 serial.tex
almost

Recently, \cite{ali2020serial} also showed that serial intervals of COVID-19 decreased in China. Analogous to our apporach, they also grouped serial intervals by the symptom onset date of infectors across 14-day periods and found that the mean serial interval decreased from 7.8 days to 2.6 days.
They attributed the decrease in serial intervals to reduction of the isolation delay period and showed that isolation delays can explain 51.5\% of the variation in serial intervals through a regression analysis.

More broadly, our study underlines the importance of carefully defining measured epidemiological time distributions.
Previous studies have shown that it matters whether generation
intervals are measured forward or backward \citep{nishiura2010time,champredon2015intrinsic,britton2019estimation};
we generalize these ideas and show that they apply to other epidemiological distributions.
Some studies during the early phases of the COVID-19 epidemics

Changes in the backward delay distributions due to changing cohort sizes are expected to be a pervasive feature of outbreak dynamics.
Recent analyses of COVID-19 epidemics have attempted to reconstruct epidemic time series from by using time-independent delay distributions from infection or symptom onset to reporting (e.g., \cite{tempvar, park2020potential,scire2020reproductive,shim2020transmission});
these studies effectively assume constant backward delay distributions.
Although such approaches may be able to roughly match the time scale of an epidemic, we show here that they are subject to bias. Dynamical approaches that model explicitly model incidence as well as reporting delays (e.g., \cite{flaxman2020estimating}) can avoid this problem, but at the cost of more complexity in estimation. 
There is a need for simple methods that can complement more explicit approaches of addressing this problem.

While our results support the use of serial-interval distributions for calculating \RR, 
they also reveal gaps in current practices in incorporating serial-interval distributions into outbreak analyses.
For example, \cite{thompson2019improved} recently emphasized the importance of using up-to-date serial-interval data for accurate estimation of time-varying reproduction numbers.
However, our results show that if changes the forward serial interval through time are not accounted for, using up-to-date serial-interval data can, in fact, exacerbate the underestimation of initial \RR.
Future studies should explore how neglecting changes in the forward serial-interval distribution can affect the estimates of \RR beyond the exponential growth phase and potentially re-assess existing estimates of \RR.

@@

While we focus here on the effect of epidemiological dynamics on serial-interval distributions, other factors, such as intervention strategies, can also affect the generation- and serial-interval distributions.
Individual-level interventions, such as case isolation, and behavioral responses directly affect individuals' ability to transmit and will shorten the forward generation- and serial-intervals.
Population-level interventions, such as social distancing, can squeeze contacts into household and induce clustering, which will in turn shorten the realized generation interval due to local depletion of the susceptible pool within households \citep{park2019inferring}.

@@

We run stochastic simulations of the renewal equation model using an individual-based model on a fully connected network (i.e., homogeneous population) based on the Gillespie algorithm that we developed earlier \citep{park2019inferring}.
First, we initialize an epidemic with $I(0)$ infected individuals (nodes) in a fully connected network of size $N$. 
For each initially infected individual, we draw number of infectious contacts from a Poisson distribution with the mean of \Ro and the corresponding generation intervals for each contact from a log-normal distribution (Table 1).
Contactees are uniformly sampled from the total population.
All contactees are sorted into event queues based on their infection time.
We update the current time to the infection time of the first person in the queue.
Then, the first person in the queue makes contacts based on the Poisson offspring distribution described earlier and their contactees are added to the sorted queue.
Whenever contactees are added to the sorted queue, we remove all duplicated contacts (but keep the first one) as well as contacts made to individuals that have already been infected.
Simulations continue until there are no more individuals in the queue.
We simulate 10 epidemics with $I(0)=10$ and $N$=40,000.

>>>>

Recently, \cite{ali2020serial} also showed that serial intervals of COVID-19 decreased in China. Analogous to our apporach, they also grouped serial intervals by the symptom onset date of infectors across 14-day periods and found that the mean serial interval decreased from 7.8 days to 2.6 days.
They attributed the decrease in serial intervals to reduction of the isolation delay; however, their regression analysis shows that isolation delays explain only 51.5\% of the variation in serial intervals (and explaining up to 72\% of the variance including other intervention measures).
Our framework provides an explanation for the remaining variation: changes in the backward incubation period during the decreasing phase of an epidemic also shorten serial intervals (even in the absence of nonpharmaceutical interventions). 
Since isolation delays and other intervention measures affect the amount of onward transmission, and therefore the distribution of realize (forward) generation intervals, it cannot explain all variation in forward serial intervals, which depend on (i) backward incubation period of the infector, (ii) the forward generation interval, and (iii) the forward incubation period of the infectee (\fref{diagram}B).

Here, we lay out a few practical principles for analyzing the serial interval data.
First, serial intervals should be cohorted based on the symptom onset date whenever possible.
Previous studies have often regarded serial intervals as an intrinsic quantity, having the same mean as the intrinsic generation interval, but their distributions are expected to change due to epidemic dynamics \citep{svensson2007note,klinkenberg2011correlation,champredon2018equivalence, britton2019estimation}.
Even when all realzied serial intervals can be observed throughout an unmitigated epidemic, we do not obtain the intrinsic serial interval distribution due to susceptible depletion (\fref{obsrR})---something true of realzied generation intervals as well \citep{park2020inferring}.
Second, aggregating serial intervals across different cohorts and epidemic periods should be avoided as they can be subject to different censoring and epidemiological biases.
Often, the amount serial interval data can be limited due to difficulties in collecting them and the aggregation cannot be avoided; in these situations, the aggregated serial-interval distributions should be interpreted with care, providing an order-of-magnitude estimates for the time scale of disease transmission.
Finally, these biases should be explicitly taken into account when inferring other epidemiological parameters (e.g., the amount of presymptomatic transmission) from serial interval data.

More broadly, our study underlines the importance of carefully defining measured epidemiological time distributions. 
Previous studies have shown that it matters whether generation
intervals are measured forward or backward \citep{nishiura2010time,champredon2015intrinsic,britton2019estimation};
we generalize these ideas and show that they apply to any other epidemiological distributions.
Some studies during the early phases of the COVID-19 epidemics have tried to correct for the backward biases, but changes in the backward delay distributions due to changing cohort sizes are expected to be a pervasive feature of outbreak dynamics.
Cohorting epidemiological delays by the primary event time can help avoid backward biases (although censoring biases can still exist) as well as detect potential changes in the distribution.

While our results support the use of serial-interval distributions for calculating the initial \RR during the exponential growth phase,  
they also reveal gaps in current practices in incorporating serial-interval distributions into outbreak analyses.

For example, \cite{thompson2019improved} recently emphasized the importance of using up-to-date serial-interval data for accurate estimation of time-varying reproduction numbers.
However, our results show that if changes the forward serial interval through time are not accounted for, using up-to-date serial-interval data can, in fact, exacerbate the underestimation of initial \RR.
Future studies should explore how neglecting changes in the forward serial-interval distribution can affect the estimates of \RR beyond the exponential growth phase and potentially re-assess existing estimates of \RR.

@@

While we focus here on the effect of epidemiological dynamics on serial-interval distributions, other factors, such as intervention strategies, can also affect the generation- and serial-interval distributions.
Individual-level interventions, such as case isolation, and behavioral responses directly affect individuals' ability to transmit and will shorten the forward generation- and serial-intervals.
Population-level interventions, such as social distancing, can squeeze contacts into household and induce clustering, which will in turn shorten the realized generation interval due to local depletion of the susceptible pool within households \citep{park2020inferring}.

@@

We run stochastic simulations of the renewal equation model using an individual-based model on a fully connected network (i.e., homogeneous population) based on the Gillespie algorithm that we developed earlier \citep{park2020inferring}.
First, we initialize an epidemic with $I(0)$ infected individuals (nodes) in a fully connected network of size $N$. 
For each initially infected individual, we draw number of infectious contacts from a Poisson distribution with the mean of \Ro and the corresponding generation intervals for each contact from a log-normal distribution (Table 1).
Contactees are uniformly sampled from the total population.
All contactees are sorted into event queues based on their infection time.
We update the current time to the infection time of the first person in the queue.
Then, the first person in the queue makes contacts based on the Poisson offspring distribution described earlier and their contactees are added to the sorted queue.
Whenever contactees are added to the sorted queue, we remove all duplicated contacts (but keep the first one) as well as contacts made to individuals that have already been infected.
Simulations continue until there are no more individuals in the queue.
We simulate 10 epidemics with $I(0)=10$ and $N$=40,000.
